High-level cognition during story listening is reflected in high-order dynamic correlations in neural activity patterns

### 2019-09-09 paper/main.tex
inputted JRM edits

\section*{Introduction}
A central goal in cognitive neuroscience is to elucidate the
\textit{neural code}: the mapping between (a) mental states or
cognitive representations and (b) neural activity patterns. One means
of testing models of the neural code is to ask how accurately that
model is able to ``translate'' neural activity patterns into known (or
hypothesized) mental states or cognitive
representations~\citep[e.g.,][]{HaxbEtal01, NormEtal06, TongPrat12,
  MitcEtal08a, KamiTong05, NishEtal11, PereEtal18, HuthEtal12,
  HuthEtal16}.  Training decoding models on different types of neural
features (Fig.~\ref{fig:patterns}a) can also help to elucidate which
specific aspects of neural activity patterns are informative about
cognition-- and, by extension, which types of neural activity patterns
might comprise the neural code.  For example, prior work has used
region of interest analyses to estimate the anatomical locations of
specific neural representations~\citep[e.g.,][]{EtzeEtal09}, or to
compare the relative contributions to the neural code of multivariate
activity patterns versus patterns of dynamic correlations between
neural activity patterns~\citep[e.g.,][]{MannEtal18, FongEtal19}.  An
emerging theme in this literature is that cognition is mediated by
dynamic interactions between brain structures~\citep{SporHone06,
  BassEtal06, Turk13, DemeEtal19, SoloEtal19, LuriEtal18, PretEtal17,
  ZouEtal19, MackEtal17}.

\begin{figure}[tp]
  \centering
  \includegraphics[width=\textwidth]{figs/patterns}
  \caption{\textbf{Neural patterns.  a. A space of neural analyses}
    Within-brain analyses are carried out within a single brain,
    whereas across-brain analyses compare neural patterns across two
    or more individuals' brains.  Univariate analyses characterize the
    activities of individual units (e.g., nodes, small networks,
    hierarchies of networks, etc.), whereas multivariate analyses
    characterize the patterns of activities across units.  Order 0
    patterns involve individual nodes; order 1 patterns involve
    node-node interactions; order 2 (and higher) patterns relate to
    interactions between homologous networks.  Each of these patterns
    may be static (e.g., averaging over time) or dynamic.
    \textbf{b. Summarizing neural patterns.}  To efficiently compute
    with complex neural patterns, it can be useful to characterize the
    patterns using summary measures.  Dimensionality reduction
    algorithms project the patterns onto lower-dimensional spaces
    whose dimensions reflect weighted combinations of the dimensions
    in the original space.  Graph measures characterize each unit's
    participation in its associated network.}
  \label{fig:patterns}
\end{figure}

@@

\section*{Methods}
Our general approach to efficiently estimating high-order dynamic
correlations comprises four general steps (Fig.~\ref{fig:methods}).
First, we derive a kernel-based approach to computing dynamic pairwise
correlations in a $T$ (timepoints) by $K$ (features) multivariate
timeseries, $\mathbf{X}_0$.  This yields a $T$ by $\mathcal{O}(K^2)$
matrix of dynamic correlations, $\mathbf{Y}_1$, where each row
comprises the upper triangle and diagonal of the correlation matrix at a single
timepoint, reshaped into a row vector (this reshaped vector is
$\left( \frac{K^2 - K}{2} + K \right)$-dimensional).  Second, we apply a dimensionality
reduction step to project the matrix of dynamic correlations back onto
a $K$-dimensional space.  This yields a $T$ by $K$ matrix,
$\mathbf{X}_1$, that reflects an approximation of the dynamic
correlations reflected in the original data.  Third, we use repeated
applications of the kernel-based dynamic correlation step to
$\mathbf{X}_n$ and the dimensionality reduction step to the resulting
$\mathbf{Y}_{n+1}$ to estimate high-order dynamic correlations.  Each
application of these steps to a $T$ by $K$ time series $\mathbf{X}_n$
yields a $T$ by $K$ matrix, $\mathbf{X}_{n+1}$, that reflects the
dynamic correlations between the columns of $\mathbf{X}_n$.  In this
way, we refer to $n$ as the \textit{order} of the timeseries, where
$\mathbf{X}_0$ (order 0) denotes the original data and $\mathbf{X}_n$
denotes (approximated) $n^\mathrm{th}$-order dynamic correlations
between the columns of $\mathbf{X}_0$.  Finally, we use a
cross-validation--based decoding approach to evaluate how well
information contained in a given order (or weighted mixture of orders)
may be used to decode relevant cognitive states.  If including a given
$\mathbf{X}_n$ in the feature set yields higher classification
accuracy on held-out data, we interpret this as evidence that the
given cognitive states are reflected in patterns of
$n^\mathrm{th}$-order correlations.  All of the code used to produce
the figures and results in this manuscript, along with links to the
corresponding datasets, may be found at
\href{https://github.com/ContextLab/timecorr-paper}{github.com/ContextLab/timecorr-paper}.
In addition, we have released a Python toolbox for computing dynamic
high-order correlations in timeseries data; our toolbox may be found
at
\href{https://timecorr.readthedocs.io/}{timecorr.readthedocs.io}. \textbf{JRM
  NOTE: CHECK LINK}

@@

\subsection*{Kernel-based approach for computing dynamic correlations}
Given a $T$ by $K$ matrix of observations, $\mathbf{X}$, we can compute the (static)
Pearson's correlation between any pair of columns, $\mathbf{X}(\cdot, i)$ and
$\mathbf{X}(\cdot, j)$ using~\citep{Pear01}:
\begin{align}
  \mathrm{corr}(\mathbf{X}(\cdot, i), \mathbf{X}(\cdot, j)) &=
                                                              \frac{\sum_{t=1}^T
                                                              \left(\mathbf{X}(t,
                                                              i)
                                                              -
                                                              \bar{\mathbf{X}}(\cdot,
                                                              i)\right)
                                                              \left(\mathbf{X}(t,
                                                              j)
                                                              -
                                                              \bar{\mathbf{X}}(\cdot, j)\right)}{\sqrt{\sum_{t=1}^T
                                                              \sigma^2_{\mathbf{X}(\cdot, i)} 
                                                              \sigma^2_{\mathbf{X}(\cdot, j)}}},~\mathrm{where}\\\label{eqn:corr}
  \bar{\mathbf{X}}(\cdot, k) &= \frac{1}{T}\sum_{t=1}^T
                               \mathbf{X}(t, k),~\mathrm{and}\\
  \sigma^2_{\mathbf{X}(\cdot, k)} &= \frac{1}{T}\sum_{t=1}^T \left( \mathbf{X}(t, k) -
                                    \bar{\mathbf{X}}(\cdot, k) \right)^2 
\end{align}
We can generalize this formula to compute time-varying correlations by
incorporating a \textit{kernel function} that takes a time $t$ as
input, and returns how much the observed data at each timepoint
$\tau \in \left[ -\infty, \infty \right]$ contributes to the estimated instantaneous
correlation at time $t$~\citep[Fig.~\ref{fig:kernels}; also see][for a
similar approach]{AlleEtal12b}.
\begin{figure}
  \centering
  \includegraphics[width=\textwidth]{figs/kernels}
  \caption{\textbf{Examples of kernel functions.} Each panel displays
    per-timepoint weights for a kernel centered at $t = 50$, evaluated
    at 100 timepoints ($\tau \in \left[1, ..., 100\right]$).
    \textbf{a. Uniform kernel.} The weights are timepoint-invariant;
    observations at all timepoints are weighted equally, and do not
    change as a function of $\tau$.  This is a special case kernel
    function that reduces dynamic correlations to static correlations.
    \textbf{b. Dirac $\delta$ kernel.} Only the observation at
    timepoint $t$ is given a non-zero weight (of 1).
    \textbf{c. Gaussian kernels.} Each kernel's weights fall off in
    time according to a Gaussian probability density function centered
    on time $t$.  Weights derived using several different example
    width parameters ($\sigma^2$) are displayed.  \textbf{d. Laplace
      kernels.}  Each kernel's weights fall off in time according to a
    Laplace probability density function centered on time $t$.
    Weights derived using several different example width parameters
    ($b$) are displayed.  \textbf{e. Mexican hat (Ricker wavelet)
      kernels.}  Each kernel's weights fall off in time according to a
    Ricker wavelet centered on time $t$.  This function highlights the
    \textit{contrasts} between local versus surrounding activity
    patterns in estimating dynamic correlations. Weights derived using
    several different example width parameters ($\sigma$) are
    displayed.}
  \label{fig:kernels}
\end{figure}

Given a kernel function $\kappa_t(\cdot)$ for timepoint $t$,
evaluated at timepoints $\tau \in \left[ 1, ..., T \right]$, we
can update the static correlation formula in Equation~\ref{eqn:corr}
to estimate the \textit{instantaneous correlation} at timepoint $t$:
\begin{align}
  \mathrm{timecorr}_{\kappa_t}\left(\mathbf{X}(\cdot, i), \mathbf{X}(\cdot, j)\right) &= \frac{\sum_{\tau=1}^T \left( \mathbf{X}(\tau, i) -
                                       \widetilde{\mathbf{X}}_{\kappa_t}(\cdot,
                                                                                        i) \right)
                                 \left( \mathbf{X}(\tau, j) -
                                        \widetilde{\mathbf{X}}_{\kappa_t}(\cdot,
                                                                                        j)\right)}
              {\sqrt{\sum_{\tau=1}^T
                                              \widetilde{\sigma}_{\kappa_t}^2(\mathbf{X}(\cdot,
                                                                                        i))
                                              \widetilde{\sigma}_{\kappa_t}^2(\mathbf{X}(\cdot, j))}},~\mathrm{where}\label{eqn:timecorr}\\
  \widetilde{\mathbf{X}}_{\kappa_t}(\cdot, k) &= \sum_{\tau=1}^T
                       \kappa_t(\tau)\mathbf{X}(\tau, k),\\
  \widetilde{\sigma}_{\kappa_t}^2(\mathbf{X}(\cdot, k)) &= \sum_{\tau=1}^T
                                                  \left(
                                                  \mathbf{X}(\tau, k) -
                            \widetilde{\mathbf{X}}_{\kappa_t}(\cdot, k) \right)^2.
\end{align}
Here
$\mathrm{timecorr}_{\kappa_t}(\mathbf{X}(\cdot, i), \mathbf{X}(\cdot,
j))$ reflects the correlation at time $t$ between columns $i$ and $j$
of $\mathbf{X}$, estimated using the kernel $\kappa_t$.  We evaluate
Equation~\ref{eqn:timecorr} in turn each pair of columns in
$\mathbf{X}$ and for kernels centered on each timepoint in the
timeseries, respectively, to obtain a $T$ by $K$ by $K$ timeseries of
dynamic correlations, $\mathbf{Y}$.  For convenience, we then reshape
the upper triangles and diagonals of each timepoint's correlation matrix into a row
vector to obtain an equivalent $T$ by $\left( \frac{K^2 - K}{2} + K \right)$ matrix.

\subsubsection*{Dynamic inter-subject functional connectivity (DISFC)}
Equation~\ref{eqn:timecorr} provides a means of taking a single
observation matrix, $\mathbf{X}_n$ and estimating the dynamic
correlations from moment to moment, $\mathbf{Y}_{n+1}$.  Suppose that
one has access to a set of multiple observation matrices that reflect
the same phenomenon.  For example, one might collect neuroimaging data
from several experimental participants, as each participant performs
the same task (or sequence of tasks).  Let $\mathbf{X}_n^1$,
$\mathbf{X}_n^2$, ..., $\mathbf{X}_n^P$ reflect the $T$ by $K$
observation matrices ($n = 0$) or reduced correlation matrices
($n > 0$) for each of $P$ participants in an experiment.  We can use
\textit{inter-subject functional connectivity}~\citep[ISFC;
][]{SimoEtal16} to compute the stimulus-driven correlations reflected
in the multi-participant dataset at a given timepoint $t$ using:
\begin{align}
\bar{\mathbf{C}}(t) = M\left(R\left(\frac{1}{2P} \sum_{p=1}^P
  Z\left(\mathbf{Y}_{n+1}^p(t)\right)^\mathrm{T} + Z\left(\mathbf{Y}_{n+1}^p(t)\right)\right)\right),\label{eqn:disfc}
\end{align}
where $M$ extracts and vectorizes the upper triangle and diagonal of a symmetric
matrix, $Z$ is the Fisher $z$-transformation~\citep{Zar10}:
\begin{align}
Z(r) = \frac{\log(1+r) - \log(1-r)}{2}
\end{align}
$R$ is the inverse of $Z$:
\begin{align}
R(z) = \frac{\exp(2z - 1)}{\exp(2z + 1)},
\end{align}
and $\mathbf{Y}_{n+1}^p(t)$ denotes the correlation matrix at timepoint $t$
(Eqn.~\ref{eqn:timecorr}) between each column of $\mathbf{X}_n^p$ and each
column of the average $\mathbf{X}_n$ from all \textit{other}
participants, $\bar{\mathbf{X}}_n^{ \backslash p}$:
\begin{align}
  \bar{\mathbf{X}}_n^{ \backslash p} = R\left(\frac{1}{P-1}\sum_{q \in
  \backslash p} Z\left( \mathbf{X}_n^q \right) \right),
\end{align}
where $\backslash p$ denotes the set of all participants other than
participant $p$. In this way, the $T$ by $\left( \frac{K^2 - K}{2} + K \right)$ DISFC
matrix $\bar{\mathbf{C}}$ provides a time-varying extension of the ISFC
approach developed by \cite{SimoEtal16}.

@@

Given a $T$ by $\left( \frac{K^2 - K}{2} + K \right)$ matrix of
dynamic correlations, $\mathbf{Y}_n$, we propose two general
approaches to computing a $T$ by $K$ low-dimensional representation of
those correlations, $\mathbf{X}_n$.  The first approach uses
dimensionality reduction algorithms to project $\mathbf{Y}_n$ onto a
$K$-dimensional space.  The second approach uses graph measures to
characterize the relative positions of each feature
($k \in \left[1, ..., K \right]$) in the network defined by the
correlation matrix at each timepoint.

@@

Applying dimensionality reduction algorithms to $\mathbf{Y}$ yields an
$\mathbf{X}$ whose columns reflect weighted combinations (or nonlinear
transformations) of the original columns of $\mathbf{Y}$.  This has
two main consequences.  First, with each repeated dimensionality
reduction, the resulting $\mathbf{X}_n$ has lower and lower fidelity
(with respect to what the ``true'' $\mathbf{Y}_n$ might have looked
like without using dimensionality reduction to maintain scalability).
In other words, computing $\mathbf{X}_n$ is a lossy operation.
Second, whereas each columns of $\mathbf{Y}_n$ may be mapped
directly onto specific pairs of columns of $\mathbf{X}_{n-1}$, the
columns of $\mathbf{X}_n$ reflect weighted combinations and/or
nonlinear transformations of the columns of $\mathbf{Y}_n$.  Many
dimensionality reduction algorithms are invertible (or approximately
invertible).  However, attempting to map a given $\mathbf{X}_n$ back
onto the original feature space of $\mathbf{X}_0$ will usually require
$\mathcal{O}(TK^{2n})$ space and therefore quickly becomes intractable
as $n$ or $K$ grow large.

@@

In general, graph measures take as input a matrix of
interactions (e.g., using the above notation, a $K$ by $K$
correlation matrix or binarized correlation matrix reconstituted from
a single timepoint's row of $\mathbf{Y}$) and return as output a set
of $K$ measures describing how each node (feature) sits within that
correlation matrix with respect to the rest of the population.  Widely
used measures include betweenness centrality~\citep[the proportion of
shortest paths between each pair of nodes in the population that
involves the given node in question; e.g.,][]{Newm05, OpsaEtal10,
  Bart04, GeisEtal08, Free77}; diversity and
dissimilarity~\citep[characterizations of how differently connected a
given node is from others in the population; e.g.,][]{Rao82, Lin09,
  RicoSzei06}; Eigenvector centrality and pagerank
centrality~\citep[measures of how influential a given node is within
the broader network; e.g.,][]{Newm08, Bona07, LohmEtal10,
  HaluEtal13}; transfer entropy and flow coefficients~\citep[a measure
of how much information is flowing from a given node to other nodes in
the network; e.g.,][]{HoneEtal07, Schr00}; $k$-coreness
centrality~\citep[a measure of the connectivity of a node within its
local sub-graph; e.g.,][]{AlvaEtal05, ChriFowl10}; within-module
degree~\citep[a measure of how many connections a node has to its
close neighbors in the network; e.g.,][]{RubiSpor10}; participation
coefficient~\citep[a measure of the diversity of a node's connections
to different sub-graphs in the network; e.g.,][]{RubiSpor10}; and
sub-graph centrality~\citep[a measure of a node's participation in all
of the network's sub-graphs; e.g.,][]{EstrRodr05}; among others.

@@

  \paragraph*{Ramping.}  We generated a third set of 10 synthetic
  datasets whose underlying correlations changed gradually over time.
  For each dataset, we constructed two \textit{anchor} correlation
  matrices using Equation~\ref{eqn:randcorr},
  $\mathbf{\Sigma}_{\mathrm{start}}$ and
  $\mathbf{\Sigma}_{\mathrm{end}}$.  For each of the 300 timepoints in
  each dataset, we drew the observations from a multivariate Gaussian
  distribution whose covariance matrix at each timepoint $t \in
  \left[0, ..., 299\right]$ was given by
  \begin{align}
    \mathbf{\Sigma}_t = \left( 1 - \frac{t}{299} \right)
    \mathbf{\Sigma}_{\mathrm{start}} + \frac{t}{299}\mathbf{\Sigma}_{\mathrm{end}}.
  \end{align}
The gradually changing correlations underlying these datasets allow us
to evaluate the recovery of dynamic correlations when each timepoint's
correlation matrix is unique (as in the random datasets), but where
the correlation dynamics are structured.

@@

After estimating an optimal set of weights, we computed a new set of
$n + 1$ $\mathbf{\Lambda}$ matrices correlating the DISFC patterns
from $\mathcal{G}_{\mathrm{train}}$ and $\mathcal{G}_{\mathrm{test}}$
at each timepoint.  We use the resulting decoding accuracy of
$\mathcal{G}_{\mathrm{test}}$ timepoints to estimate how informative
the set of neural features containing up to $n^\mathrm{th}$ order
correlations were.

>>>>

\section*{Introduction}
A central goal in cognitive neuroscience is to elucidate the
\textit{neural code}: the mapping between (a) mental states or
cognitive representations and (b) neural activity patterns. One means
of testing models of the neural code is to ask how accurately that
model is able to ``translate'' neural activity patterns into known (or
hypothesized) mental states or cognitive
representations~\citep[e.g.,][]{HaxbEtal01, NormEtal06, TongPrat12,
  MitcEtal08a, KamiTong05, NishEtal11, PereEtal18, HuthEtal12,
  HuthEtal16}.  Training decoding models on different types of neural
features (Fig.~\ref{fig:patterns}a) can also help to elucidate which
specific aspects of neural activity patterns are informative about
cognition-- and, by extension, which types of neural activity patterns
might comprise the neural code.  For example, prior work has used
region of interest analyses to estimate the anatomical locations of
specific neural representations~\citep[e.g.,][]{EtzeEtal09}, or to
compare the relative contributions to the neural code of multivariate
activity patterns versus dynamic correlations between neural activity
patterns~\citep[e.g.,][]{MannEtal18, FongEtal19}.  An emerging theme
in this literature is that cognition is mediated by dynamic
interactions between brain structures~\citep{SporHone06, BassEtal06,
  Turk13, DemeEtal19, SoloEtal19, LuriEtal18, PretEtal17, ZouEtal19,
  MackEtal17}.

\begin{figure}[tp]
  \centering
  \includegraphics[width=\textwidth]{figs/patterns}
  \caption{\textbf{Neural patterns.  a. A space of neural analyses}
    Within-brain analyses are carried out within a single brain,
    whereas across-brain analyses compare neural patterns across two
    or more individuals' brains.  Univariate analyses characterize the
    activities of individual units (e.g., nodes, small networks,
    hierarchies of networks, etc.), whereas multivariate analyses
    characterize the patterns of activities across units.  Order 0
    patterns involve individual nodes; order 1 patterns involve
    node-node interactions; order 2 (and higher) patterns relate to
    interactions between homologous networks.  Each of these patterns
    may be static (e.g., averaging over time) or dynamic.
    \textbf{b. Summarizing neural patterns.}  To efficiently compute
    with complex neural patterns, it can be useful to characterize the
    patterns using summary measures.  Dimensionality reduction
    algorithms project the patterns onto lower-dimensional spaces
    whose dimensions reflect weighted combinations or non-linear
    transformations of the dimensions in the original space.  Graph
    measures characterize each unit's participation in its associated
    network.}
  \label{fig:patterns}
\end{figure}

@@

\section*{Methods}
Our general approach to efficiently estimating high-order dynamic
correlations comprises four general steps (Fig.~\ref{fig:methods}).
First, we derive a kernel-based approach to computing dynamic pairwise
correlations in a $T$ (timepoints) by $K$ (features) multivariate
timeseries, $\mathbf{X}_0$.  This yields a $T$ by $\mathcal{O}(K^2)$
matrix of dynamic correlations, $\mathbf{Y}_1$, where each row
comprises the upper triangle and diagonal of the correlation matrix at a single
timepoint, reshaped into a row vector (this reshaped vector is
$\left( \frac{K^2 - K}{2} + K \right)$-dimensional).  Second, we apply a dimensionality
reduction step to project the matrix of dynamic correlations back onto
a $K$-dimensional space.  This yields a $T$ by $K$ matrix,
$\mathbf{X}_1$, that reflects an approximation of the dynamic
correlations reflected in the original data.  Third, we use repeated
applications of the kernel-based dynamic correlation step to
$\mathbf{X}_n$ and the dimensionality reduction step to the resulting
$\mathbf{Y}_{n+1}$ to estimate high-order dynamic correlations.  Each
application of these steps to a $T$ by $K$ time series $\mathbf{X}_n$
yields a $T$ by $K$ matrix, $\mathbf{X}_{n+1}$, that reflects the
dynamic correlations between the columns of $\mathbf{X}_n$.  In this
way, we refer to $n$ as the \textit{order} of the timeseries, where
$\mathbf{X}_0$ (order 0) denotes the original data and $\mathbf{X}_n$
denotes (approximated) $n^\mathrm{th}$-order dynamic correlations
between the columns of $\mathbf{X}_0$.  Finally, we use a
cross-validation--based decoding approach to evaluate how well
information contained in a given order (or weighted mixture of orders)
may be used to decode relevant cognitive states.  If including a given
$\mathbf{X}_n$ in the feature set yields higher classification
accuracy on held-out data, we interpret this as evidence that the
given cognitive states are reflected in patterns of
$n^\mathrm{th}$-order correlations.  All of the code used to produce
the figures and results in this manuscript, along with links to the
corresponding datasets, may be found at
\href{https://github.com/ContextLab/timecorr-paper}{github.com/ContextLab/timecorr-paper}.
In addition, we have released a Python toolbox for computing dynamic
high-order correlations in timeseries data; our toolbox may be found
at
\href{https://timecorr.readthedocs.io/}{timecorr.readthedocs.io}.

@@

\subsection*{Kernel-based approach for computing dynamic correlations}
Given a $T$ by $K$ matrix of observations, $\mathbf{X}$, we can compute the (static)
Pearson's correlation between any pair of columns, $\mathbf{X}(\cdot, i)$ and
$\mathbf{X}(\cdot, j)$ using~\citep{Pear01}:
\begin{align}
  \mathrm{corr}(\mathbf{X}(\cdot, i), \mathbf{X}(\cdot, j)) &=
                                                              \frac{\sum_{t=1}^T
                                                              \left(\mathbf{X}(t,
                                                              i)
                                                              -
                                                              \bar{\mathbf{X}}(\cdot,
                                                              i)\right)
                                                              \left(\mathbf{X}(t,
                                                              j)
                                                              -
                                                              \bar{\mathbf{X}}(\cdot, j)\right)}{\sqrt{\sum_{t=1}^T
                                                              \sigma^2_{\mathbf{X}(\cdot, i)} 
                                                              \sigma^2_{\mathbf{X}(\cdot, j)}}},~\mathrm{where}\label{eqn:corr}\\
  \bar{\mathbf{X}}(\cdot, k) &= \frac{1}{T}\sum_{t=1}^T
                               \mathbf{X}(t, k),~\mathrm{and}\\
  \sigma^2_{\mathbf{X}(\cdot, k)} &= \frac{1}{T}\sum_{t=1}^T \left( \mathbf{X}(t, k) -
                                    \bar{\mathbf{X}}(\cdot, k) \right)^2 
\end{align}
We can generalize this formula to compute time-varying correlations by
incorporating a \textit{kernel function} that takes a time $t$ as
input, and returns how much the observed data at each timepoint
$\tau \in \left[ -\infty, \infty \right]$ contributes to the estimated instantaneous
correlation at time $t$~\citep[Fig.~\ref{fig:kernels}; also see][for a
similar approach]{AlleEtal12b}.
\begin{figure}
  \centering
  \includegraphics[width=\textwidth]{figs/kernels}
  \caption{\textbf{Examples of kernel functions.} Each panel displays
    per-timepoint weights for a kernel centered at $t = 50$, evaluated
    at 100 timepoints ($\tau \in \left[1, ..., 100\right]$).
    \textbf{a. Uniform kernel.} The weights are timepoint-invariant;
    observations at all timepoints are weighted equally, and do not
    change as a function of $\tau$.  This is a special case kernel
    function that reduces dynamic correlations to static correlations.
    \textbf{b. Dirac $\delta$ kernel.} Only the observation at
    timepoint $t$ is given a non-zero weight (of 1).
    \textbf{c. Gaussian kernels.} Each kernel's weights fall off in
    time according to a Gaussian probability density function centered
    on time $t$.  Weights derived using several different example
    width parameters ($\sigma^2$) are displayed.  \textbf{d. Laplace
      kernels.}  Each kernel's weights fall off in time according to a
    Laplace probability density function centered on time $t$.
    Weights derived using several different example width parameters
    ($b$) are displayed.  \textbf{e. Mexican hat (Ricker wavelet)
      kernels.}  Each kernel's weights fall off in time according to a
    Ricker wavelet centered on time $t$.  This function highlights the
    \textit{contrasts} between local versus surrounding activity
    patterns in estimating dynamic correlations. Weights derived using
    several different example width parameters ($\sigma$) are
    displayed.}
  \label{fig:kernels}
\end{figure}

Given a kernel function $\kappa_t(\cdot)$ for timepoint $t$,
evaluated at timepoints $\tau \in \left[ 1, ..., T \right]$, we
can update the static correlation formula in Equation~\ref{eqn:corr}
to estimate the \textit{instantaneous correlation} at timepoint $t$:
\begin{align}
  \mathrm{timecorr}_{\kappa_t}\left(\mathbf{X}(\cdot, i), \mathbf{X}(\cdot, j)\right) &= \frac{\sum_{\tau=1}^T \left( \mathbf{X}(\tau, i) -
                                       \widetilde{\mathbf{X}}_{\kappa_t}(\cdot,
                                                                                        i) \right)
                                 \left( \mathbf{X}(\tau, j) -
                                        \widetilde{\mathbf{X}}_{\kappa_t}(\cdot,
                                                                                        j)\right)}
              {\sqrt{\sum_{\tau=1}^T
                                              \widetilde{\sigma}_{\kappa_t}^2(\mathbf{X}(\cdot,
                                                                                        i))
                                              \widetilde{\sigma}_{\kappa_t}^2(\mathbf{X}(\cdot, j))}},~\mathrm{where}\label{eqn:timecorr}\\
  \widetilde{\mathbf{X}}_{\kappa_t}(\cdot, k) &= \sum_{\tau=1}^T
                       \kappa_t(\tau)\mathbf{X}(\tau, k),\\
  \widetilde{\sigma}_{\kappa_t}^2(\mathbf{X}(\cdot, k)) &= \sum_{\tau=1}^T
                                                  \left(
                                                  \mathbf{X}(\tau, k) -
                            \widetilde{\mathbf{X}}_{\kappa_t}(\cdot, k) \right)^2.
\end{align}
Here
$\mathrm{timecorr}_{\kappa_t}(\mathbf{X}(\cdot, i), \mathbf{X}(\cdot,
j))$ reflects the correlation at time $t$ between columns $i$ and $j$
of $\mathbf{X}$, estimated using the kernel $\kappa_t$.  We evaluate
Equation~\ref{eqn:timecorr} in turn for each pair of columns in
$\mathbf{X}$ and for kernels centered on each timepoint in the
timeseries, respectively, to obtain a $T$ by $K$ by $K$ timeseries of
dynamic correlations, $\mathbf{Y}$.  For convenience, we then reshape
the upper triangles and diagonals of each timepoint's correlation matrix into a row
vector to obtain an equivalent $T$ by $\left( \frac{K^2 - K}{2} + K \right)$ matrix.

\subsubsection*{Dynamic inter-subject functional connectivity (DISFC)}
Equation~\ref{eqn:timecorr} provides a means of taking a single
observation matrix, $\mathbf{X}_n$ and estimating the dynamic
correlations from moment to moment, $\mathbf{Y}_{n+1}$.  Suppose that
one has access to a set of multiple observation matrices that reflect
the same phenomenon.  For example, one might collect neuroimaging data
from several experimental participants, as each participant performs
the same task (or sequence of tasks).  Let $\mathbf{X}_n^1$,
$\mathbf{X}_n^2$, ..., $\mathbf{X}_n^P$ reflect the $T$ by $K$
observation matrices ($n = 0$) or reduced correlation matrices
($n > 0$) for each of $P$ participants in an experiment.  We can use
\textit{inter-subject functional connectivity}~\citep[ISFC;
][]{SimoEtal16} to compute the stimulus-driven correlations reflected
in the multi-participant dataset at a given timepoint $t$ using:
\begin{align}
\bar{\mathbf{C}}(t) = M\left(R\left(\frac{1}{2P} \sum_{p=1}^P
  Z\left(\mathbf{Y}_{n+1}^p(t)\right)^\top + Z\left(\mathbf{Y}_{n+1}^p(t)\right)\right)\right),\label{eqn:disfc}
\end{align}
where $M$ extracts and vectorizes the upper triangle and diagonal of a symmetric
matrix, $Z$ is the Fisher $z$-transformation~\citep{Zar10}:
\begin{align}
Z(r) = \frac{\log(1+r) - \log(1-r)}{2},
\end{align}
$R$ is the inverse of $Z$:
\begin{align}
R(z) = \frac{\exp(2z - 1)}{\exp(2z + 1)},
\end{align}
and $\mathbf{Y}_{n+1}^p(t)$ denotes the correlation matrix at timepoint $t$
(Eqn.~\ref{eqn:timecorr}) between each column of $\mathbf{X}_n^p$ and each
column of the average $\mathbf{X}_n$ from all \textit{other}
participants, $\bar{\mathbf{X}}_n^{ \backslash p}$:
\begin{align}
  \bar{\mathbf{X}}_n^{ \backslash p} = \frac{1}{P-1}\sum_{q \in
  \backslash p} \mathbf{X}_n^q,
\end{align}
where $\backslash p$ denotes the set of all participants other than
participant $p$. In this way, the $T$ by $\left( \frac{K^2 - K}{2} + K \right)$ DISFC
matrix $\bar{\mathbf{C}}$ provides a time-varying extension of the ISFC
approach developed by \cite{SimoEtal16}.

@@

Given a $T$ by $\left( \frac{K^2 - K}{2} + K \right)$ matrix of
$n^\mathrm{th}$-order dynamic correlations, $\mathbf{Y}_n$, we propose
two general approaches to computing a $T$ by $K$ low-dimensional
representation of those correlations, $\mathbf{X}_n$.  The first
approach uses dimensionality reduction algorithms to project
$\mathbf{Y}_n$ onto a $K$-dimensional space.  The second approach uses
graph measures to characterize the relative positions of each feature
($k \in \left[1, ..., K \right]$) in the network defined by the
correlation matrix at each timepoint.

@@

Applying dimensionality reduction algorithms to $\mathbf{Y}$ yields an
$\mathbf{X}$ whose columns reflect weighted combinations (or nonlinear
transformations) of the original columns of $\mathbf{Y}$.  This has
two main consequences.  First, with each repeated dimensionality
reduction, the resulting $\mathbf{X}_n$ has lower and lower fidelity
(with respect to what the ``true'' $\mathbf{Y}_n$ might have looked
like without using dimensionality reduction to maintain scalability).
In other words, computing $\mathbf{X}_n$ is a lossy operation.
Second, whereas each column of $\mathbf{Y}_n$ may be mapped
directly onto specific pairs of columns of $\mathbf{X}_{n-1}$, the
columns of $\mathbf{X}_n$ reflect weighted combinations and/or
nonlinear transformations of the columns of $\mathbf{Y}_n$.  Many
dimensionality reduction algorithms are invertible (or approximately
invertible).  However, attempting to map a given $\mathbf{X}_n$ back
onto the original feature space of $\mathbf{X}_0$ will usually require
$\mathcal{O}(TK^{2n})$ space and therefore becomes intractable
as $n$ or $K$ grow large.

@@

In general, graph measures take as input a matrix of
interactions (e.g., using the above notation, a $K$ by $K$
correlation matrix or binarized correlation matrix reconstituted from
a single timepoint's row of $\mathbf{Y}$), and return as output a set
of $K$ measures describing how each node (feature) sits within that
correlation matrix with respect to the rest of the population.  Widely
used measures include betweenness centrality~\citep[the proportion of
shortest paths between each pair of nodes in the population that
involves the given node in question; e.g.,][]{Newm05, OpsaEtal10,
  Bart04, GeisEtal08, Free77}; diversity and
dissimilarity~\citep[characterizations of how differently connected a
given node is from others in the population; e.g.,][]{Rao82, Lin09,
  RicoSzei06}; eigenvector centrality and pagerank
centrality~\citep[measures of how influential a given node is within
the broader network; e.g.,][]{Newm08, Bona07, LohmEtal10,
  HaluEtal13}; transfer entropy and flow coefficients~\citep[a measure
of how much information is flowing from a given node to other nodes in
the network; e.g.,][]{HoneEtal07, Schr00}; $k$-coreness
centrality~\citep[a measure of the connectivity of a node within its
local subgraph; e.g.,][]{AlvaEtal05, ChriFowl10}; within-module
degree~\citep[a measure of how many connections a node has to its
close neighbors in the network; e.g.,][]{RubiSpor10}; participation
coefficient~\citep[a measure of the diversity of a node's connections
to different subgraphs in the network; e.g.,][]{RubiSpor10}; and
subgraph centrality~\citep[a measure of a node's participation in all
of the network's subgraphs; e.g.,][]{EstrRodr05}; among others.

@@

  \paragraph*{Ramping.}  We generated a third set of 10 synthetic
  datasets whose underlying correlations changed gradually over time.
  For each dataset, we constructed two \textit{anchor} covariance
  matrices using Equation~\ref{eqn:randcorr},
  $\mathbf{\Sigma}_{\mathrm{start}}$ and
  $\mathbf{\Sigma}_{\mathrm{end}}$.  For each of the 300 timepoints in
  each dataset, we drew the observations from a multivariate Gaussian
  distribution whose covariance matrix at each timepoint $t \in
  \left[0, ..., 299\right]$ was given by
  \begin{align}
    \mathbf{\Sigma}_t = \left( 1 - \frac{t}{299} \right)
    \mathbf{\Sigma}_{\mathrm{start}} + \frac{t}{299}\mathbf{\Sigma}_{\mathrm{end}}.
  \end{align}
The gradually changing correlations underlying these datasets allow us
to evaluate the recovery of dynamic correlations when each timepoint's
correlation matrix is unique (as in the random datasets), but where
the correlation dynamics are structured.

@@

After estimating an optimal set of weights, we computed a new set of
$n + 1$ $\mathbf{\Lambda}$ matrices correlating the DISFC patterns
from $\mathcal{G}_{\mathrm{train}}$ and $\mathcal{G}_{\mathrm{test}}$
at each timepoint.  We use the resulting decoding accuracy of
$\mathcal{G}_{\mathrm{test}}$ timepoints (using the weights in
$\phi_{0, 1, ..., n}$ to average the $\Lambda$ matrices) to estimate how informative
the set of neural features containing up to $n^\mathrm{th}$ order
correlations were.
